---
title: "Approaches to Missing and Censored Data in Biomarker Research"
output: 
  html_document:
    code_folding: show
    theme:
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---

```{r setup, include=FALSE}
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.showtext = TRUE, comment = NA) 

```

```{r include=FALSE}
# packageurl <- "https://cran.r-project.org/src/contrib/Archive/gsl/gsl_2.1-8.tar.gz"
# install.packages(packageurl, repos = NULL, type = "source")
# install.packages("clusterMI")
library(clusterMI)
library(inaparc)
library(mclust)
library(missMDA)
library(FactoMineR)
library(ClusterR)
library(Rfast)
library(parallel)
library(inaparc)
library(kpodclustr)
library(ISLR)

source(file = "functions.R")
```

```{r theme, include=FALSE}

##theme for plots 
library(ggplot2)
theme_publication <- theme_bw() +
  theme(
    # Titles
    plot.title = element_text(size = 9, hjust = 0.5),
    axis.title = element_text(size = 8, ),
    axis.text  = element_text(size = 8),
    
    # Legend
    legend.title = element_text(size = 7),
    legend.text  = element_text(size = 6),
    legend.position = "right",   
    legend.key.size = unit(0.6, "cm"), 
    
    # Margin
    plot.margin = margin(3, 3, 3, 3)
  )

```
## 4)	The case of omics data

```{r data, include=FALSE}
set.seed(1234)
data("NCI60", package = "ISLR")
nci.lab <- NCI60$labs
nci.data <- NCI60$data
k <- length(unique(nci.lab))

# adding 30% of missing values MCAR 
xxna<- clusterMI::prodna(nci.data, pct = 0.3)
```

### Principal components - section 4.a.i - figure 15.1

```{r, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

## number of components (time consumming)
# res.ncp <- estim_ncpPCA(xxna,
#                     ncp.min = 5,
#                     ncp.max = 20,
#                     scale = TRUE,
#                     method.cv = "Kfold",
#                     nbsim = 10,
#                     verbose = TRUE)
# ncp <- 16
# res.imp <- imputePCA(xxna, ncp = ncp, scale = TRUE)
# res.pca <- PCA(res.imp$completeObs, graph = FALSE)
#save(res.ncp, res.imp, res.pca, file = "pca_results.RData")
load("data/pca_results.RData")

## Figure 15.1
col <- gray.colors(k)
names(col) <- unique(nci.lab)
col <- col[nci.lab]
names(col) <- NULL

#png("iterativepca.png", res=300,width=2000, height=2000)
plot.new()
plot(res.pca,
     choix = "ind",
     col.ind = col,
     graph.type="classic",
     label = "none",
     cex=1.5)
B <- array2DF(by(res.pca$ind$coord[,1:2],INDICES =  col,FUN = colmeans), simplify = TRUE)
points(B[,2:3],col=B[,1],pch=3,cex=3,lwd=2)
legend("topright", pch = 3,cex = .8, lwd = 2,legend = unique(nci.lab), ncol = 2, bty = "n", col = col)
#dev.off()
```

```{r, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
#png("pca.png", res=300,width=2000, height=2000)
res.pca.full <- PCA(nci.data, graph=FALSE)
plot.new()
plot(res.pca.full,
     choix = "ind",
     col.ind = col,
     graph.type="classic",
     label = "none",
     cex=1.5)
B <- array2DF(by(res.pca.full$ind$coord[,1:2],INDICES =  col,FUN = colmeans), simplify = TRUE)
points(B[,2:3],col=B[,1],pch=3,cex=3,lwd=2)
legend("topright", pch = 3,cex = .8, lwd = 2,legend = unique(nci.lab), ncol = 2, bty = "n", col = col)
#dev.off()
```

### Clustering - section 4.a.ii - figure 15.2


```{r, include=TRUE}

## direct
#res.kpod <- kpod(X = as.matrix(xxna), k = k)
#save(res.kpod, file = "kpod_results.RData")
load("data/kpod_results.RData")

## Figure 15.2
res.ca <- CA(table(res.kpod$cluster,nci.lab), graph=FALSE)

#png("ca_clustering.png", res=300,width=2000, height=2000)
print(plot(res.ca, graph.type = "classic",cex=1.2,col.row = "black", col.col = "darkgrey",xlim = c(-1.2,3.5)))
#dev.off()
```

### Subspace clustering - section 4.a.iii - figure 15.3

```{r, include=TRUE}

# MI

# res.imp <- MIPCA(xxna, ncp = 16, scale = TRUE, nboot = 20,verbose = TRUE)
 # res.ana <- lapply(res.imp$res.MI,
 #                       FUN = KMeans_rcpp ,
 #                       clusters = k,num_init=10)
# 
#save(res.imp, res.ana, file = "imp_results.RData")
load("data/imp_results.RData")

### extract the set of partitions (as list)
res.ana<- lapply(res.ana, "[[", "clusters")
### pooling by NMF
res.pool<- fastnmf(res.ana, nb.clust = k,method.init = c("BOK","kmeans"))
part.km <- res.pool$best$clust

# rkpod 
# res.rkpod <- rkpod(xxna=xxna,
#                    k=k,
#                    q=6,
#                    nnodes=7,
#                    method.init.rkpod = rep("MIPCA",7),
#                    ncp=16)
# 
# save(res.rkpod, file = "rkpod_results.RData")
load("data/rkpod_results.RData")

## Figure 15.3
#png("rkpod_map.png", res=300,width=2000, height=2000)
print(plot.rkpod(res.rkpod))
#dev.off()
```

## 5)	The case of unique biomarkers

### Data simulation

We first simulate data based on the distribution of biomarkers (BMK) to better evaluate the impact of missing values on our statistical outcomes. This simulation, which includes fitting a logistic regression, serves as an illustrative example to demonstrate how missing data can influence efficacy related analyses. 

```{r}
library(ggplot2)
library(ggpubr)

set.seed( 454)

N <- 150

age <- rnorm(n = N, mean = 50, sd = 20)
gender <-  rbinom(n = N, size = 1, p=0.5)
mutation <-  rbinom(n = N, size = 1, p=0.2)
biomarker <- age*0.2 + 6*gender + 6*mutation + rnorm(N, mean = 0, sd = 4)

stat <-  -2 + biomarker*0.2
proba <- plogis(stat)
ORR <- rbinom(N,1,proba)

datasim <- data.frame(age = age, gender = as.factor(gender), mutation = mutation,
                      biomarker = biomarker, 
                      proba = proba, 
                      ORR = ORR)
```


### Data exploration {.tabset .tabset-pills}

#### Biomarker according to ORR

```{r}
corrplot <- ggboxplot(datasim, "ORR", "biomarker",
          color = "ORR", palette =c("#00AFBB", "#E7B800"),
          add = "jitter", shape = "ORR") + ggtitle("Biomarker distribution according to ORR")
corrplot

```


#### Biomarker according to age

```{r}
ggscatter(datasim, x = "age", y = "biomarker",
          color = "black", shape = 21, size = 3, # Points color, shape and size
          add = "reg.line",  # Add regressin line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE, # Add confidence interval
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          cor.coeff.args = list(method = "pearson", label.x = 30, label.sep = "\n")) +
            ggtitle("Biomarker according to age")

```

#### Biomarker according to gender

```{r}
ggboxplot(datasim, "gender", "biomarker",
          color = "gender", palette =c("#00AFBB", "#E7B800"),
          add = "jitter", shape = "gender")+
            ggtitle("Biomarker according to gender")

```

#### Biomarker according to mutation

```{r}
ggboxplot(datasim, "mutation", "biomarker",
          color = "mutation", palette =c("#00AFBB", "#E7B800"),
          add = "jitter", shape = "mutation")+
            ggtitle("Biomarker according to mutation")

```

### Logistic regression fit figure 15.4


```{r}

LOD <- quantile(datasim$biomarker, probs = 0.4)

modelfull <- glm(ORR ~ biomarker, family = "binomial", data = datasim)
#summary(modelfull)$coef

logfit <- ggplot(data = datasim, aes(biomarker, ORR)) +
  geom_point(alpha = .15, aes(color = as.factor(ORR))) +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
    geom_vline(xintercept = LOD, linetype="dotdash", color = "gray") + 
  ggtitle("Logistic regression model fit") +
  xlab("Biomarker") +
  ylab("Probability of the event") +
    theme_bw() 

logfit

ggsave(filename = "logfit.png", plot = logfit)

# logfitbw <- ggplot(data = datasim, aes(biomarker, ORR)) +
#   geom_point(aes(color = as.factor(ORR))) +
#   geom_smooth(method = "glm", color = "black", method.args = list(family = "binomial")) +
#     geom_vline(xintercept = LOD, linetype="dotdash", color = "gray") + 
#   ggtitle("Logistic regression model fit") +
#   xlab("Biomarker") +
#   ylab("Probability of the event") +
#     theme_bw() +
#   scale_color_manual(
#     name = "Response",
#     values = c("0" = "black", "1" = "gray")
#   ) 

#logfitbw

#ggsave(filename = "logfitbw.png", plot = logfitbw, height = 8, width = 8)

```


### Adding missing data and single imputation

```{r}
library(gridExtra)

datasim$biomarkerNA  <- datasim$biomarker
datasim$biomarkerNA[datasim$biomarkerNA < LOD] <- NA
datasim$cens  <- is.na(datasim$biomarkerNA)

datasim$biomarkerlod  <- datasim$biomarker
datasim$biomarkerlod[datasim$biomarkerlod < LOD] <- LOD

datasim$biomarkerlodsqrt2  <- datasim$biomarker
datasim$biomarkerlodsqrt2[datasim$biomarkerlodsqrt2 < LOD] <- LOD/sqrt(2)

datasim$biomarkerlod2  <- datasim$biomarker
datasim$biomarkerlod2[datasim$biomarkerlod2 < LOD] <- LOD/2

datasim$biomarker0  <- datasim$biomarker
datasim$biomarker0[datasim$biomarker0 < LOD] <- 0

plot0 <-  ggplot(datasim, aes(x=biomarker0)) + 
  geom_density() + ggtitle("Imputation by 0")# + xlim(-2, 10)

plotlod2 <-  ggplot(datasim, aes(x=biomarkerlod2)) + 
  geom_density() +  ggtitle("Imputation by LOD/2") #xlim(0, 10) +

plotlodsqrt2 <-  ggplot(datasim, aes(x=biomarkerlodsqrt2)) + 
  geom_density()  + ggtitle("Imputation by LOD/sqrt(2)") #+ xlim(0, 10)

plotlod <-  ggplot(datasim, aes(x=biomarkerlod)) + 
  geom_density() + ggtitle("Imputation by LOD") #xlim(0, 10) + 

plotfull <-  ggplot(datasim, aes(x=biomarker)) + 
  geom_density()  + ggtitle("Full data") #+ xlim(-10, 10) + ylim(0, 0.4)

plotcomp <-  ggplot(datasim, aes(x=biomarkerNA)) + 
  geom_density()  + ggtitle("Complete Case") #+ xlim(-10, 10) + ylim(0, 0.4)

grid.arrange(plotfull, plotcomp, plotlod, 
             plotlodsqrt2, plotlod2, plot0,
             nrow = 2)

```


### Multiple imputation from section 5.c - figure 15.5


```{r}
library(qgcomp)
library(mice)
library(dplyr)
library(tidyr)
#Imputation based on linear left censored model based on age + gender + mutation as covariate
#10 imputation in this example - could be 5 as discussed in the book

impdat <- mice::mice(data = datasim[, c("age", "gender", "mutation", "biomarkerNA")],
              method = c("", "", "", "leftcenslognorm"),
              lod=c(NA, NA, NA, LOD), debug=FALSE, m=10, print = FALSE)

mires <- as.data.frame(matrix(unlist(impdat$imp), ncol = 10))
colnames(mires) <- paste("imp", 1:10, sep ="_")

mireslong <- mires %>%
  pivot_longer(cols = starts_with("imp"),, names_to = "type", values_to = "biomarker")

mireslong <- rbind(mireslong, 
                   data.frame(type = rep("biomarkercens", length = N), biomarker = datasim$biomarkerNA)) %>% 
  mutate(value = case_when(type == "biomarkercens" ~ "observed", 
                           .default  = "imputed"))
  
densitymi <- ggplot(mireslong, aes(x=biomarker, fill=type)) +
  geom_density(alpha=0.4, color = "gray") +
  scale_fill_manual(values = c("darkcyan", rep("goldenrod", 10))) +
   theme_bw() 


#densitymi

ggsave(filename = "densitymi.png", plot = densitymi)


##Black and white figures for the book
bw = 1
densitymibw <- ggplot(mireslong, aes(x = biomarker, fill = type)) +
  geom_density(alpha = 0.7, bw = bw) +
  # couleurs pour chaque imputation + observed
  scale_fill_manual(
    values = c("biomarkercens" = "black", rep("gray", 10)),
    guide = "none"   # enlève la légende liée à 'type'
  ) +
  # ajoute une légende simple basée sur 'value'
  geom_density(aes(color = value), alpha = 0.7,  bw = bw) +
  scale_color_manual(
    name = "values",
    values = c("observed" = "black", "imputed" = "gray")
  ) +
  ggtitle("Comparison of Observed and Imputed Biomarker Distributions")+
  theme_publication

#densitymibw

ggsave(filename = "densitymibw.png", plot = densitymibw)


boxplotmibw <- ggplot(mireslong, aes(y = biomarker, x = type, color = value)) +
  geom_boxplot(fill = "white") +
  geom_point() + 
  # couleurs pour chaque imputation + observed
  scale_color_manual(
    name = "values",
    values = c("observed" = "black", "imputed" = "gray")
  ) + 
  geom_hline(yintercept = LOD, color = "black", linetype = "dashed")+
  ggtitle("Comparison of Observed and Imputed Biomarker Distributions")+
  ylab("Biomarker values") + xlab("") + 
  theme_publication +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

boxplotmibw

ggsave(filename = "boxplot.png", plot = boxplotmibw, width = 7, height = 5)


```

### 5.a	Descriptive analysis of left censored biomarkers

i-Kaplan Meier 

```{r}
library(NADA)
library(censReg)

##Kaplan Meier for left censored:
km <- NADA::cenfit(datasim$biomarker, datasim$cens)

##Display mean +/- se
descriptive_km <- mean(km)
descriptive_km

```


ii-ROS 

```{r}
#regression on order statistics:
ros <- NADA::cenros(datasim$biomarker, is.na(datasim$cens))

#Imputed values
ros_values <- ros$modeled 

#Mean & SD & SE
mean_ros <- mean(ros_values)
se_ros <- sd(ros_values) / sqrt(length(ros_values))

# 95% confidence interval
ci_lower_ros <- mean_ros - 1.96 * se_ros
ci_upper_ros <- mean_ros + 1.96 * se_ros

# Combine results
descriptive_ros <- c(mean_ros, se_ros, ci_lower_ros, ci_upper_ros)
names(descriptive_ros) <- names(descriptive_km)
descriptive_ros

```



Tobit

```{r}
tobit <- censReg::censReg(datasim$biomarkerlod~1, left = LOD)

# Estimateur (moyenne)
mean_tobit <- coef(tobit)[1]

# Erreur standard associée
se_tobit <- sqrt(diag(vcov(tobit)))[1]

# Intervalle de confiance à 95%
ci_lower_tobit <- mean_tobit - 1.96 * se_tobit
ci_upper_tobit <- mean_tobit + 1.96 * se_tobit

# Combine results
descriptive_tobit <- c(mean_tobit, se_tobit, ci_lower_tobit, ci_upper_tobit)
names(descriptive_tobit) <- names(descriptive_km)
descriptive_tobit

```



Comparison with full model and single imputation

```{r}
#Full model 
mean_full <- mean(datasim$biomarker)
se_full <- sd(datasim$biomarker) / sqrt(length(datasim$biomarker))

#Complete Case
mean_cc <- mean(datasim$biomarkerNA, na.rm = TRUE)
se_cc <- sd(datasim$biomarkerNA, na.rm = TRUE) / sqrt(sum(!is.na(datasim$biomarkerNA)))

#LOD
mean_lod <- mean(datasim$biomarkerlod, na.rm = TRUE)
se_lod <- sd(datasim$biomarkerlod, na.rm = TRUE) / sqrt(length(datasim$biomarkerlod))

#LOD/2
mean_lod2 <- mean(datasim$biomarkerlod2)
se_lod2 <- sd(datasim$biomarkerlod2) / sqrt(length(datasim$biomarkerlod2))


#LOD/sqrt(2)
mean_lodsqrt2 <- mean(datasim$biomarkerlodsqrt2)
se_lodsqrt2 <- sd(datasim$biomarkerlodsqrt2) / sqrt(length(datasim$biomarkerlodsqrt2))

#0
mean_0 <- mean(datasim$biomarker0)
se_0 <- sd(datasim$biomarker0) / sqrt(length(datasim$biomarker0))



# creating a data frame and adding the 95% confidence interval
descriptive <- data.frame(mean = c(mean_full, mean_cc, mean_lod, mean_lod2, mean_lodsqrt2, mean_0), 
                   se = c(se_full, se_cc, se_lod, se_lod2, se_lodsqrt2, se_0))
                    
descriptive$ci_lower <- descriptive$mean - 1.96 * descriptive$se
descriptive$ci_upper <- descriptive$mean + 1.96 * descriptive$se


descriptive <- as.data.frame(rbind(descriptive, 
                     descriptive_km, 
                   descriptive_ros, 
                   descriptive_tobit
                   ))

descriptive$label <- c("Full data", "Complete Case", "LOD imputation", 
                        "LOD/2 imputation", "LOD/sqrt(2) imputation", "0 imputation", 
                       "Kaplan Meier", "ROS", "Tobit")
descriptive$index <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)

##Forest plot to compare results
forestplotmean <- ggplot(descriptive, aes(y = index, x = mean)) +
  geom_point(shape = 18, size = 5) +  
  geom_errorbarh(aes(xmin = ci_lower , xmax = ci_upper), height = 0.25) +
  geom_vline(xintercept = mean_full, color = "red", linetype = "dashed", cex = 1, alpha = 0.5) +
  scale_y_continuous(name = "", breaks=1:9, labels = descriptive$label, trans = "reverse") +
  xlab("Mean (95% CI)") + 
  ylab(" ") + 
  theme_bw() +
  theme(panel.border = element_blank(),
        #panel.background = element_blank(),
        #panel.grid.major = element_blank(), 
        #panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x.bottom = element_text(size = 12, colour = "black"),
        axis.title.x = element_text(size = 12, colour = "black")) +
  ggtitle("mean estimation according to missing data handling")

forestplotmean


ggsave(filename = "forestplotmean2.png", plot = forestplotmean, width = 12, height = 8 )

```

### 6)	Incomplete data reporting and sensitivity assessment - fig 15.6

```{r}
##fit with full data: 
modelfull <- glm(ORR ~ biomarker, family = "binomial", data = datasim)
##fit with complete case: 
modelcomp <- glm(ORR ~ biomarkerNA, family = "binomial", data = datasim)
##fit with LOD: 
modellod <- glm(ORR ~ biomarkerlod, family = "binomial", data = datasim)
##fit with LOD/2: 
modellod2 <- glm(ORR ~ biomarkerlod2, family = "binomial", data = datasim)
##fit with LOD/sqrt2: 
modellodsqrt2 <- glm(ORR ~ biomarkerlodsqrt2, family = "binomial", data = datasim)
##fit with 0: 
model0 <- glm(ORR ~ biomarker0, family = "binomial", data = datasim)
##fit MI: 
fit.imp.log <- impdat %>%
  with(glm(ORR ~ biomarkerNA, family = "binomial"))

modelMI <- pool(fit.imp.log)
```


```{r}
##Combine results
logisticdata <- as.data.frame(rbind(summary(modelfull)$coef[2, ], 
                      summary(modelcomp)$coef[2, ], 
                      summary(modellod)$coef[2, ], 
                      summary(modellod2)$coef[2, ],
                      summary(modellodsqrt2)$coef[2, ],
                      summary(model0)$coef[2, ], 
                      summary(modelMI)[2, c("estimate", "std.error", "statistic", "p.value")]))

#Renaiming because of different convention between mice and glm
rownames(logisticdata) <- c("full", "complete", "lod", "lod2", "lodsqrt2", "0", "MI")
colnames(logisticdata) <- names(summary(modelfull)$coef[2, ])

logisticdata$LL <- logisticdata$Estimate - qt(0.975,df=N-1)*logisticdata$`Std. Error`/sqrt(N)
logisticdata$UL <- logisticdata$Estimate + qt(0.975,df=N-1)*logisticdata$`Std. Error`/sqrt(N)
logisticdata$IC <- paste(round(logisticdata$LL, 2),
                         round(logisticdata$UL, 2), sep = ", ")
logisticdata$label <- c("Full data", "Complete Case", "LOD imputation", 
                        "LOD/2 imputation", "LOD/sqrt(2) imputation", "0 imputation", "Multiple imputation")
logisticdata$index <- c(1, 2, 3, 4, 5, 6, 7)

logisticdata

##Forest plot
forestplot <- ggplot(logisticdata, aes(y = index, x = Estimate)) +
  geom_point(shape = 18, size = 5, color = "black") +  
  geom_errorbarh(aes(xmin = LL, xmax = UL), height = 0.25, color = "black") +
  geom_vline(xintercept = summary(modelfull)$coef[2, 1], color = "red", linetype = "dashed", cex = 1, alpha = 0.5) +
  scale_y_continuous(name = "", breaks=1:7, labels = logisticdata$label, trans = "reverse") +
  xlab("Odds Ratio (95% CI)") + 
  ylab(" ") + 
  theme_bw() +
  # theme(panel.border = element_blank(),
  #       panel.background = element_blank(),
  #       panel.grid.major = element_blank(), 
  #       panel.grid.minor = element_blank(), 
  #       axis.line = element_line(colour = "black"),
  #       axis.text.y = element_text(size = 12, colour = "black"),
  #       axis.text.x.bottom = element_text(size = 12, colour = "black"),
  #       axis.title.x = element_text(size = 12, colour = "black")) +
  ggtitle("Logistic fit according to missing data handling")

forestplot

ggsave(filename = "forestplot.png", plot = forestplot)

```


```{r}
library(cowplot)

forestplotbw <- ggplot(logisticdata, aes(y = index, x = Estimate)) +
  geom_point(shape = 18, size = 5, color = "black") +  
  geom_errorbarh(aes(xmin = LL, xmax = UL), height = 0.25, color = "black") +
  geom_vline(xintercept = summary(modelfull)$coef[2, 1], color = "black", linetype = "dashed", cex = 1, alpha = 0.5) +
  scale_y_continuous(name = "", breaks=1:7, labels = logisticdata$label, trans = "reverse") +
  xlab("Odds Ratio (95% CI)") + 
  ylab(" ") + 
  theme_publication +
  ggtitle("Logistic fit according to missing data handling")

forestplotbw

forestplotmeanbw <- ggplot(descriptive, aes(y = index, x = mean)) +
  geom_point(shape = 18, size = 5, color = "black") +  
  geom_errorbarh(aes(xmin = ci_lower , xmax = ci_upper), height = 0.25, color = "black") +
  geom_vline(xintercept = mean_full, color = "black", linetype = "dashed", cex = 1, alpha = 0.5) +
  scale_y_continuous(name = "", breaks=1:9, labels = descriptive$label, trans = "reverse") +
  xlab("Mean (95% CI)") + 
  ylab(" ") + 
   theme_publication +
  ggtitle("Mean estimation according to missing data handling")

forestplotmeanbw

# Combining 2 plots
forest_combined_plot <- plot_grid(forestplotmeanbw, forestplotbw,
                           labels = c("A", "B"),   
                           ncol = 2, align = "h")  

# Titre global
title <- ggdraw() + 
  draw_label("Sensitivity of Mean and Logistic Regression Estimates to Imputation Techniques", fontface = 'bold', x = 0.5, hjust = 0.5, size = 10)

final_plot <- plot_grid(title, forest_combined_plot, ncol = 1, rel_heights = c(0.1, 1))
plotforchapter <- plot_grid(forest_combined_plot, ncol = 1, rel_heights = c(0.1, 1))

# Affichage
final_plot
#plotforchapter
# Sauvegarde
ggsave("forest_combined_plot.png", plotforchapter, width = 10, height = 4, dpi = 300)

```



